---
title: 'Reproducible Research: Peer Assessment 1'
author: "Wenwen Liu"
output: html_document
---

### Loading and processing the data
1. Load the data
```{r load}
activity <- read.csv("activity.csv", na.strings = "NA")
```
2. Transform the date format
```{r date}
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```
### Histogram of the total number of steps taken each day
1.Calculate the total number of steps
```{r message=FALSE, warning=FALSE}
library(dplyr)
perday <- activity%>%
  group_by(date)%>%
  summarize(sumsteps = sum(steps))
```
2.Make a histogram of the total number of steps taken each day
```{r message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(perday, aes(x = sumsteps)) + 
  geom_histogram(binwidth = 2000, color = "black",fill = "#d25959") +
  ylim(0,20) +
  labs(title = "Histogram of Total Steps Taken Per Day (Oct 1st to 12th, 2012 )") +
  labs(x = "Total steps per day",y = "Frequency")
```

3.Calculate and report the mean and median of the total number of steps taken per day
```{r}
mday <- perday %>%
  summarize(meansteps = mean(sumsteps, na.rm=TRUE), 
            mediansteps = median(sumsteps, na.rm=TRUE))
mean<-as.integer(round(mday[1],digits=0));median<-mday[2]
print(mday)
```
   The mean and median of the toal number of steps taken per day are `r mean` and `r median`, respectively.

### The average daily active pattern
1. Time series plot of the 5-minute interval and the average number of steps taken
```{r}
perminute<-activity%>%
  group_by(interval)%>%
  summarize(meansteps=mean(steps, na.rm=TRUE))
ggplot(perminute, aes(x=interval, y=meansteps))+geom_line(color="#d25959",size=0.7)+
  labs(title="Average Activity Pattern Per Day")+labs(x="Time Interval",y="Steps")
```

2. Find the 5-minute interval that contains the maxium average number of steps
```{r}
result<-perminute[which.max(perminute$meansteps),]
```
   The interval `r result[1]` contains the maxium average steps at `r result[2]`

### Imputing missing values
1.Calculate and report the total number of missing values in the dataset.
```{r}
totalna<-sum(is.na(activity$steps))
```
   There are in total `r totalna` missing values in the original dataset.
2.Use the mean for that specific 5-minute interval in place of missing values.
```{r}
activity1<-activity
for(i in 1:dim(activity)[1]){
  if(is.na(activity$steps[i])){
    activity1$steps[i]=perminute$meansteps[which(perminute$interval==activity$interval[i])]
  }
}
```
3.Make a histogram of the total number of steps taken each day and report the mean and median.
```{r}
library(dplyr)
perday1<-activity1%>%
  group_by(date)%>%
  summarize(sumsteps=sum(steps))
library(ggplot2)
ggplot(perday1, aes(x=sumsteps)) + geom_histogram(binwidth =2000, color="black",fill="#d25959")+
  labs(title="Histogram of Total Steps Taken Per Day, Imputed (Oct 1st to 12th, 2012)")+
  labs(x="Total steps per day",y="Frequency")
mday1<-perday1%>%
  summarize(meansteps=mean(sumsteps, na.rm=TRUE), mediansteps=median(sumsteps, na.rm=TRUE))
  mean1<-as.integer(round(mday[1]));median1<-as.integer(round(mday[2]))
```
The mean and median total number of steps taken per day after imputation are `r mean1` and `r median1`. They are the same with the previous means and median acquired before imputation. 

## Activity patterns between weekdays and weekends
1.Create a new factor variable in the dataset with two levels – “weekday” and “weekend”.
```{r}
wd<-c()
for(i in 1:dim(activity1)[1]){
  if(as.factor(weekdays(activity1$date[i]))=="Saturday"|as.factor(weekdays(activity1$date[i]))=="Sunday"){
    wd<-c(wd, "weekend")
  }
  else{
    wd<-c(wd, "weekday")
  }
}
activity1<-cbind(activity1, wd)
```
2.Time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.
```{r}
perweekday<-activity1%>%
  group_by(wd, interval)%>%
  summarize(meansteps=mean(steps, na.rm=TRUE))
p <- ggplot(perweekday, aes(x=interval, y=meansteps,color=wd)) + geom_line(size=0.7)
p + facet_grid(wd~.,space="free")+theme(legend.title=element_blank(), plot.title=element_text(hjust = 0.5))+ 
  labs(x="Time Interval",y="Number of Steps")+theme(legend.position="none")+
  labs(title="Average Daily Steps by Day Type")
```


